### 2018-04-06 abstract.tex
Outline abstract/Intro. Add rough title
Draft to Committee

\abstract{
An ABSTRACT describing your amazing work goes here.
}

>>>>

\abstract{
The direct detection of gravitational waves from the inspiral and merger of compact object binaries by the Advanced LIGO and Virgo interferometers has ushered in an exciting new era of astronomy. Analysis of this gravitational wave data provides fundamental insight into GR in a strong gravity regime not normally accessible and allows the detection of binary systems not visible via electromagnetic observations. However, truly maximizing the science gains from these detections requires the joint detection of a coincident electromagnetic counterpart. Doing so will provde new insight into the envionment and host galaxy of the merger, an accurate determination of distance and energy scales, and insight into the hydrodynamics of the merger. The most promising counterpart for this task is the ``kilonovae," an optical/NIR transient powered by the radioactive decay of heavy $r$-process elements synthesized in the merger ejecta.

In this thesis, I present a series of studies that culminate in the first joint detection of gravitational waves and electromagnetic radiation from the merger of a binary neutron star system. First, I will present studies of how to design and execute observational strategies to detect optical counterparts of interest and reject contaminating false-positives. Using both simulated and empirical observations I show that kilonovae can be well separated from other transients on the basis of their timescales and color, and that the rate of contaminating transients will be low.

Second, I will present our efforts to follow-up gravitational wave events with the Dark Energy Camera (DECam). In particular, I will focus on our follow-up of the first binary neutron star merger detected by Advanced LIGO and Virgo, GW170817. Our DECam program was able to make an independent discovery of the optical counterpart to GW170817. I will present my modeling of the broadband optical/NIR photometry that shows this optical emission is consistent with expectations for a kilonova and that the amount of material ejected during the merger is sufficient to suggest that binary neutron star mergers can be a dominant site of cosmic $r$-process nucleosynthesis.
}

### 2018-04-06 acknowledge.tex
Fix Refs. Chapter 5 Bulk Formatting
Draft to Committee

%Chapter 2 Stuff
We thank Ryan Chornock, Maria Drout, Wen-fai Fong, Ryan Foley, Daniel Kasen, Brian Metzger, Armin Rest, and Ken Shen for helpful discussions and providing model data during the course of this analysis. P.S.C. is grateful for support provided by the NSF through the Graduate Research Fellowship Program, grant DGE1144152.

%Chapter 3 Stuff
The Berger Time-Domain Group at Harvard is supported in part by the NSF through grants AST-1411763 and AST-1714498, and by NASA through grants NNX15AE50G and NNX16AC22G. P.S.C. is grateful for support provided by the NSF
through the Graduate Research Fellowship Program, grant DGE1144152. The UCSC group is supported in part by NSF grant AST--1518052, the Gordon \& Betty Moore Foundation, and from fellowships from the Alfred P.\ Sloan Foundation and the David and Lucile Packard Foundation to R.J.F.

This work has made use of data from the European Space Agency (ESA)
mission \gaia (\url{https://www.cosmos.esa.int/gaia}), processed by
the \gaia Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the \gaia Multilateral Agreement.
This publication makes use of data products from the Two Micron All Sky Survey,
which is a joint project of the University of Massachusetts and the Infrared Processing
and Analysis Center/California Institute of Technology, funded by the National Aeronautics
and Space Administration and the National Science Foundation. This publication makes use
of data products from the Wide-field Infrared Survey Explorer, which is a joint project of the
University of California, Los Angeles, and the Jet Propulsion Laboratory/California Institute of
Technology, funded by the National Aeronautics and Space Administration.

% Chapter 4 Stuff
P.S.C. is grateful for support provided by the NSF
through the Graduate Research Fellowship Program, grant DGE1144152.
R.J.F.\ gratefully acknowledges support from NSF grant AST--1518052 and
the Alfred P.\ Sloan Foundation. D.E.H.\ was supported by NSF CAREER grant
PHY-1151836. He also acknowledges support from the Kavli Institute for
Cosmological Physics at the University of Chicago through NSF grant
PHY-1125897 as well as an endowment from the Kavli Foundation.

This research uses services or data provided by the NOAO Science
Archive. NOAO is operated by the Association of Universities for
Research in Astronomy (AURA), Inc. under a cooperative agreement with
the National Science Foundation. The computations in this paper were
run on the Odyssey cluster supported by the FAS Division of Science,
Research Computing Group at Harvard University. This research has made
use of the NASA/IPAC Extragalactic Database (NED) which is operated by
the Jet Propulsion Laboratory, California Institute of Technology,
under contract with the National Aeronautics and Space Administration.
Light curve data for PS15cdi were obtained from The Open Supernova
Catalog \citep{Guillochon+17}. Some of the results in this paper have been
derived using the HEALPix package \citep{Gorski+05}.

Funding for the DES Projects has been provided by the DOE and NSF (USA),
MEC/MICINN/ MINECO (Spain), STFC (UK), HEFCE (UK). NCSA (UIUC),
KICP (U. Chicago), CCAPP (Ohio State), MIFPA (Texas A\&M), CNPQ, FAPERJ,
FINEP (Brazil), DFG (Germany) and the Collaborating Institutions in the Dark Energy Survey.

The Collaborating Institutions are Argonne Lab, UC Santa Cruz,
University of Cambridge, CIEMAT-Madrid, University of Chicago, University College London,
DES-Brazil Consortium, University of Edinburgh, ETH Z{\"u}rich, Fermilab,
University of Illinois, ICE (IEEC-CSIC), IFAE Barcelona, Lawrence Berkeley Lab,
LMU M{\"u}nchen and the associated Excellence Cluster Universe,
University of Michigan, NOAO, University of Nottingham, Ohio State University, University of
Pennsylvania, University of Portsmouth, SLAC National Lab, Stanford University,
University of Sussex, Texas A\&M University, and the OzDES Membership Consortium.

The DES Data Management System is supported by the NSF under
Grant Number AST-1138766. The DES participants from Spanish institutions are partially
supported by MINECO under grants AYA2012-39559, ESP2013-48274, FPA2013-47986,
and Centro de Excelencia Severo Ochoa SEV-2012-0234. Research leading
to these results has received funding from the ERC under the EU's 7$^{\rm th}$
Framework Programme including grants ERC 240672, 291329 and 306478.

% Chapter 5
The Berger Time-Domain Group at Harvard is supported in part by the NSF through grants AST-1411763 and AST-1714498, and by NASA through grants NNX15AE50G and NNX16AC22G. PSC is grateful for support provided by the NSF through the Graduate Research Fellowship Program, grant DGE 1144152. VAV acknowledges support by the National Science Foundation through a Graduate Research Fellowship. DAB is supposed by NSF award PHY-1707954. We thank the University of Copenhagen, DARK Cosmology Centre, and the Niels Bohr International Academy for hosting R.J.F.\ during the discovery of GW170817/SSS17a, where he was participating in the Kavli Summer Program in Astrophysics, ``Astrophysics with gravitational wave detections."  This program was supported by the Kavli Foundation, Danish National Research Foundation, the Niels Bohr International Academy, and the DARK Cosmology Centre. The UCSC group is supported in part by NSF grant AST--1518052, the Gordon \& Betty Moore Foundation, the Heising-Simons Foundation, generous donations from many individuals through a UCSC Giving Day grant, and from fellowships from the Alfred P.\ Sloan Foundation and the David and Lucile Packard Foundation to R.J.F. EB acknowledges financial support from the European Research Council (ERC-StG-335936, CLUSTERS). DK \& EQ were funded in part by the Gordon and Betty Moore Foundation through Grant GBMF5076.

We are grateful for the heroic efforts of the entire staff at Gemini-South to continue obtaining observations of this target in evening twilight at high airmass as the object was setting.

This project used data obtained with the Dark Energy Camera (DECam), which was constructed by the Dark Energy Survey (DES) collaboration. Funding for the DES Projects has been provided by the DOE and NSF (USA), MISE (Spain), STFC (UK), HEFCE (UK), NCSA (UIUC), KICP (U. Chicago), CCAPP (Ohio State), MIFPA (Texas A\&M), CNPQ, FAPERJ, FINEP (Brazil), MINECO (Spain), DFG (Germany) and the collaborating institutions in the Dark Energy Survey, which are Argonne Lab, UC Santa Cruz, University of Cambridge, CIEMAT-Madrid, University of Chicago, University College London, DES-Brazil Consortium, University of Edinburgh, ETH Z{\"u}rich, Fermilab, University of Illinois, ICE (IEEC-CSIC), IFAE Barcelona, Lawrence Berkeley Lab, LMU M{\"u}nchen and the associated Excellence Cluster Universe, University of Michigan, NOAO, University of Nottingham, Ohio State University, University of Pennsylvania, University of Portsmouth, SLAC National Lab, Stanford University, University of Sussex, and Texas A\&M University.

Based in part on observations at Cerro Tololo Inter-American Observatory, National Optical Astronomy Observatory (NOAO Prop. ID: 2017B-0110, PI: E. Berger), which is operated by the Association of Universities for Research in Astronomy (AURA) under a cooperative agreement with the National Science Foundation

Based in part on observations obtained at the Gemini Observatory (Program IDs GS-2017B-Q-8 and GS-2017B-DD-4; PI: Chornock), which is operated by the Association of Universities for Research in Astronomy, Inc., under a cooperative agreement with the NSF on behalf of the Gemini partnership: the National Science Foundation (United States), the National Research Council (Canada), CONICYT (Chile), Ministerio de Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina), and Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil).

>>>>

Acknowledgments forthcoming. There are far too many people to thank...

%The Berger Time-Domain Group at Harvard is supported in part by the NSF through grants AST-1411763 and AST-1714498, and by NASA through grants NNX15AE50G and NNX16AC22G. P.S.C. is grateful for support provided by the NSF through the Graduate Research Fellowship Program, grant DGE1144152. V.A.V. acknowledges support by the National Science Foundation through a Graduate Research Fellowship. EB acknowledges financial support from the European Research Council (ERC-StG-335936, CLUSTERS). DK \& EQ were funded in part by the Gordon and Betty Moore Foundation through Grant GBMF5076. DAB is supposed by NSF award PHY-1707954.
%
%The UCSC group is supported in part by NSF grant AST--1518052, the Gordon \& Betty Moore Foundation, the Heising-Simons Foundation, generous donations from many individuals through a UCSC Giving Day grant, and from fellowships from the Alfred P.\ Sloan Foundation and the David and Lucile Packard Foundation to R.J.F. R.J.F.\ gratefully acknowledges support from NSF grant AST--1518052 and the Alfred P.\ Sloan Foundation. D.E.H.\ was supported by NSF CAREER grant PHY-1151836. He also acknowledges support from the Kavli Institute for Cosmological Physics at the University of Chicago through NSF grant PHY-1125897 as well as an endowment from the Kavli Foundation.
%
%R.J.F.\ thanks the University of Copenhagen, DARK Cosmology Centre, and the Niels Bohr International Academy for hosting during the discovery of GW170817/SSS17a, where he was participating in the Kavli Summer Program in Astrophysics, ``Astrophysics with gravitational wave detections."  This program was supported by the Kavli Foundation, Danish National Research Foundation, the Niels Bohr International Academy, and the DARK Cosmology Centre.
%
%This thesis uses services or data provided by the NOAO Science Archive. NOAO is operated by the Association of Universities for Research in Astronomy (AURA), Inc. under a cooperative agreement with the National Science Foundation. The computations in this thesis were run on the Odyssey cluster supported by the FAS Division of Science, Research Computing Group at Harvard University. This thesis has made use of the NASA/IPAC Extragalactic Database (NED) which is operated by the Jet Propulsion Laboratory, California Institute of Technology, under contract with the National Aeronautics and Space Administration. Some of the results in this thesis have been derived using the HEALPix package \citep{Gorski+05}.
%
%The work presented in Chapter 2 relied upon Ryan Chornock, Maria Drout, Wen-fai Fong, Ryan Foley, Daniel Kasen, Brian Metzger, Armin Rest, and Ken Shen for helpful discussions and providing model data during the course of the analysis.
%
%This thesis has made use of data from the European Space Agency (ESA) mission \gaia\ (\url{https://www.cosmos.esa.int/gaia}), processed by the \gaia Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC has been provided by national institutions, in particular the institutions participating in the \gaia Multilateral Agreement. This thesis makes use of data products from the Two Micron All Sky Survey, which is a joint project of the University of Massachusetts and the Infrared Processing and Analysis Center/California Institute of Technology, funded by the National Aeronautics and Space Administration and the National Science Foundation. This thesis makes use of data products from the Wide-field Infrared Survey Explorer, which is a joint project of the University of California, Los Angeles, and the Jet Propulsion Laboratory/California Institute of Technology, funded by the National Aeronautics and Space Administration.
%
%The light curve data for PS15cdi used in Chapter 4 were obtained from The Open Supernova
%Catalog \citep{Guillochon+17a}.
%
%This thesis used data obtained with the Dark Energy Camera (DECam), which was constructed by the Dark Energy Survey (DES) collaboration. Funding for the DES Projects has been provided by the U.S. Department of Energy, the U.S. National Science Foundation, the Ministry of Science and Education of Spain, the Science and Technology FacilitiesCouncil of the United Kingdom, the Higher Education Funding Council for England, the National Center for Supercomputing Applications at the University of Illinois at Urbana-Champaign, the Kavli Institute of Cosmological Physics at the University of Chicago, the Center for Cosmology and Astro-Particle Physics at the Ohio State University, the Mitchell Institute for Fundamental Physics and Astronomy at Texas A\&M University, Financiadora de Estudos e Projetos, Funda{\c c}{\~a}o Carlos Chagas Filho de Amparo {\`a} Pesquisa do Estado do Rio de Janeiro, Conselho Nacional de Desenvolvimento Cient{\'i}fico e Tecnol{\'o}gico and the Minist{\'e}rio da Ci{\^e}ncia, Tecnologia e Inova{\c c}{\~a}o, the Deutsche Forschungsgemeinschaft, and the Collaborating Institutions in the Dark Energy Survey.
%
%The Collaborating Institutions are Argonne National Laboratory, the University of California at Santa Cruz, the University of Cambridge, Centro de Investigaciones Energ{\'e}ticas, Medioambientales y Tecnol{\'o}gicas-Madrid, the University of Chicago, University College London, the DES-Brazil Consortium, the University of Edinburgh, the Eidgen{\"o}ssische Technische Hochschule (ETH) Z{\"u}rich,  Fermi National Accelerator Laboratory, the University of Illinois at Urbana-Champaign, the Institut de Ci{\`e}ncies de l'Espai (IEEC/CSIC), the Institut de F{\'i}sica d'Altes Energies, Lawrence Berkeley National Laboratory, the Ludwig-Maximilians Universit{\"a}t M{\"u}nchen and the associated Excellence Cluster Universe, the University of Michigan, the National Optical Astronomy Observatory, the University of Nottingham, The Ohio State University, the OzDES Membership Consortium, the University of Pennsylvania, the University of Portsmouth, SLAC National Accelerator Laboratory, Stanford University, the University of Sussex, and Texas A\&M University.
%
%The DES Data Management System is supported by the NSF under Grant Number AST-1138766. The DES participants from Spanish institutions are partially supported by MINECO under grants AYA2012-39559, ESP2013-48274, FPA2013-47986, and Centro de Excelencia Severo Ochoa SEV-2012-0234. Research leading to these results has received funding from the ERC under the EU's 7$^{\rm th}$ Framework Programme including grants ERC 240672, 291329 and 306478.
%
%This thesis is based in part on observations at Cerro Tololo Inter-American Observatory, National Optical Astronomy Observatory (NOAO Prop. ID: 2017B-0110, PI: E. Berger), which is operated by the Association of Universities for Research in Astronomy (AURA) under a cooperative agreement with the National Science Foundation
%
%This thesis is based in part on observations obtained at the Gemini Observatory (Program IDs GS-2017B-Q-8 and GS-2017B-DD-4; PI: Chornock), which is operated by the Association of Universities for Research in Astronomy, Inc., under a cooperative agreement with the NSF on behalf of the Gemini partnership: the National Science Foundation (United States), the National Research Council (Canada), CONICYT (Chile), Ministerio de Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina), and Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil).
%
%We are grateful for the heroic efforts of the entire staff at Gemini-South to continue obtaining observations of GW170817 in evening twilight at high airmass as the object was setting.

### 2018-04-06 chapter4/ms2.tex
Draft to Committee

\section{Introduction}
\label{sec:ch4_intro}

The Advanced Laser Interferometer Gravitational-Wave Observatory
(LIGO) is designed to detect the final inspiral and
merger of compact object binaries comprised of neutron stars (NS)
and/or stellar-mass black holes (BH) \citep{LIGOMainRef}. 
The first LIGO observing run (designated O1)
commenced on 18 September 2015 with the ability to detect binary
neutron star (BNS) mergers to an average distance of $\approx 75$ Mpc, a
forty-fold increase in volume relative to the previous generation of
ground-based GW detectors \citep{Martynov+16}. On 2015 September 14 LIGO detected 
the first GW event ever observed, GW150914 \citep{LIGOGW150914}.

@@

A second high-significance GW event, designated GW151226, was
discovered by LIGO on 2015 December 26 at 03:38:53 UT \citep{LIGOGW151226}. 
This event was also due to the inspiral and merger of a BBH system, consisting 
of $14.2^{+8.3}_{-3.7}$~M$_\odot$ and $7.5^{+2.3}_{-2.3}$~M$_\odot$ black holes at a luminosity 
distance of $d_L = 440^{+180}_{-190}$~Mpc \citep{LIGOGW151226}.
The initial localization was provided as a probability sky map via a private 
GCN circular 38 hours after the GW detection \citep{SingerPrice16}. We initiated
optical follow-up observations with DECam 10 hours later on 2015 December 28,
and imaged a 28.8 deg$^2$ region in the $i$ and $z$ bands during several epochs. Here we report 
the results of this search. In \cref{sec:ch4_obs} we discuss the
observations and data analysis procedures. In \cref{sec:ch4_analysis} 
we present our search methodology for potential counterparts to GW151226, 
and the results of this search. We summarize our conclusions in \cref{sec:ch4_conc}. We perform 
cosmological calculations assuming $H_0 = 67.8$~km  s$^{-1}$ Mpc$^{-1}$,
$\Omega_{\lambda} = 0.69$, and $\Omega_m = 0.31$ \citep{Planck2016}. 
Magnitudes are reported in the AB system.

\clearpage
\section{Observations and Data Reduction} 
\label{sec:ch4_obs}

GW151226 was detected on 2015 December 26 at 03:38:53 UT by a 
Compact Binary Coalescence (CBC) search pipeline \citep{LIGOGW151226}. The CBC
pipeline operates by matching the strain data against waveform
templates and is sensitive to mergers containing NS and/or BH. 
The initial sky map was generated by the {\tt BAYESTAR} algorithm
and released 38 hours after the GW detection. {\tt BAYESTAR} is a Bayesian 
algorithm that generates a localization sky map based on the parameter
estimation from the CBC pipeline \citep{Singer+14,SingerPrice16}. The sky area contained within the
initial 50\% and 90\% contours was 430 deg$^2$ and 1340 deg$^2$,
respectively. A sky map generated by the {\tt LALInference}
algorithm, which computes the localization using Bayesian forward-modeling 
of the signal morphology \citep{Veitch+15}, was released on 2016 January 15 UT, 
after our DECam observations had been concluded. The {\tt LALInference} sky map
is slightly narrower than the sky map from {\tt BAYESTAR} with 50\% and 90\% contours 
of 362 deg$^2$ and 1238 deg$^2$, respectively.

@@

\begin{deluxetable}{lrccccccc}
\singlespace
\tabletypesize{\footnotesize}
\tablecolumns{9}
\tablewidth{0pt}
\tablecaption{Summary of DECam Observations of GW151226
\label{tab:ch4_datatable}}
\tablehead{	
\colhead{Visit}    &  
\colhead{UT}       & 
\colhead{$\Delta t$\tablenotemark{a}} &
\colhead{$\langle$PSF$_i$$\rangle$}    &
\colhead{$\langle$PSF$_z$$\rangle$}    &
\colhead{$\langle$airmass$\rangle$}  &
\colhead{$\langle$depth$_i\rangle$}   &
\colhead{$\langle$depth$_z\rangle$}  &
\colhead{$A_{\mathrm{eff}}$\tablenotemark{b}} \\
\colhead{ } & 
\colhead{ } &
\colhead{(days)} &
\colhead{(arcsec)} &
\colhead{(arcsec)} &
\colhead{ } &
\colhead{(mag)} &
\colhead{(mag)} &
\colhead{(deg$^2$)} 
}
\startdata
Epoch 1 & 2015-12-28.11 & 1.96 & 0.97 & 0.99 & 1.95 & 22.39 & 22.23 & 14.4 \\      %  6 hexes 
   & 2015-12-29.11 & 2.96 & 1.00 & 0.97 & 1.78 & 22.57 & 22.46 & 14.4 \\      %  6 hexes

Epoch 2 & 2016-01-01.06 & 5.91 & 0.95 & 0.90 & 1.57 & 21.37 & 21.06 & 28.8 \\      % 12 hexes

Epoch 3 & 2016-01-08.11 & 12.96 & 1.68 & 1.62 & 2.15 & 22.09 & 21.70 & 24.0 \\      % 10 hexes
    & 2016-01-09.11 & 13.96 & 1.17 & 1.12 & 1.80 & 22.44 & 22.17 &  4.8 \\      %  2 hexes

Epoch 4 & 2016-01-18.03 & 22.88 & 1.21 & 1.20 & 1.48 & 22.00 & 22.01 & 12.0 \\      %  5 hexes
   & 2016-01-19.01 & 23.86 & 1.29 & 1.25 & 1.71 & 21.86 & 21.90 & 16.8 \\ %  7 hexes
\enddata
\tablecomments{Summary of our DECam follow-up observations of
GW151226. The PSF, airmass, and depth are the average values across
all observations on that date. The reported depth corresponds to the mean 
$5\sigma$ point source detection in the coadded search images.}
\tablenotetext{a}{Time elapsed between the GW trigger time and the time of the first image.} 
\tablenotetext{b}{The effective area corresponds to 12 DECam pointings 
taking into account that $\approx 20\%$ of the 3
deg$^2$ field of view of DECam is lost due to chip gaps (10\%), 3
dead CCDs (5\%, \citealt{Diehl+14}), and masked edge pixels (5\%).}
\end{deluxetable}

@@

\section{Search for an Optical Counterpart}
\label{sec:ch4_analysis}

The primary focus of our search is a fast-fading transient. 
While the merger of a BBH system is not expected
to produce an EM counterpart, it is informative to consider the
possibility of optical emission due to the presence of some matter in
the system. As a generic example, we consider the behavior of a
transient such as a short gamma-ray burst (SGRB) with a typical beaming-corrected energy of
$E_j \approx 10^{49}$ erg and an opening angle of $\theta_j\approx
10^\circ$ \citep{Berger2014,Fong+15}. If viewed far off-axis $(\theta_{\rm
obs}\gtrsim 4\theta_j)$ the optical emission will reach peak
brightness after several days, but at the distance of GW151226
($\approx440$ Mpc, \citealt{LIGOGW151226}), the peak brightness will be $i\approx 26$ mag (see Figure 5 of
\citealt{MetzgerBerger12}), well beyond our detection limit. If the source
is observed moderately off-axis or on-axis $(\theta_{\rm obs} \lesssim
2\theta_j)$, then the light curve will decline throughout our
observations, roughly as $F_\nu\propto t^{-1}$, and will be detectable
at $i\approx$ 21--22 mag in our first observation (see Figures 3 and 4
of \citealt{MetzgerBerger12}). We can apply a similar argument to the
behavior of a more isotropic (and non-relativistic) outflow given that
any material ejected in a BBH merger is likely to have a low mass and
the outflow will thus become optically thin early, leading to fading
optical emission. Based on this line of reasoning, we search our data
for steadily declining transients.

We identify relevant candidates in the data using the following
selection criteria with the forced photometry from {\tt photpipe}. 
Unless otherwise noted these criteria are applied to the $i$-band 
data due to the greater depth in those observations.

\begin{enumerate}

\item We require non-negative or consistent with zero 
(i.e., within 2$\sigma$ of zero) 
$i$- and $z$-band fluxes in the difference photometry
across all epochs to eliminate any sources that re-brighten in the 
fourth (template) epoch. This provides an initial sample of 602 candidates.

>>>>

\section{Introduction}
\label{sec:ch4_intro}
The Advanced Laser Interferometer Gravitational-Wave Observatory (LIGO) is designed to detect the final inspiral and merger of compact object binaries comprised of neutron stars (NS) and/or stellar-mass black holes (BH) \citep{LIGOMainRef}.  The first LIGO observing run (designated O1) commenced on 18 September 2015 with the ability to detect binary neutron star (BNS) mergers to an average distance of $\approx 75$ Mpc, a forty-fold increase in volume relative to the previous generation of ground-based GW detectors \citep{Martynov+16}. On 2015 September 14 LIGO detected  the first GW event ever observed, GW150914 \citep{LIGOGW150914}.

@@

A second high-significance GW event, designated GW151226, was discovered by LIGO on 2015 December 26 at 03:38:53 UT \citep{LIGOGW151226}.  This event was also due to the inspiral and merger of a BBH system, consisting  of $14.2^{+8.3}_{-3.7}$~M$_\odot$ and $7.5^{+2.3}_{-2.3}$~M$_\odot$ black holes at a luminosity  distance of $d_L = 440^{+180}_{-190}$~Mpc \citep{LIGOGW151226}. The initial localization was provided as a probability sky map via a private  GCN circular 38 hours after the GW detection \citep{SingerPrice16}. We initiated optical follow-up observations with DECam 10 hours later on 2015 December 28, and imaged a 28.8 deg$^2$ region in the $i$ and $z$ bands during several epochs. Here we report  the results of this search.

In \cref{sec:ch4_obs} we discuss the observations and data analysis procedures. In \cref{sec:ch4_analysis}  we present our search methodology for potential counterparts to GW151226,  and the results of this search. We summarize our conclusions in \cref{sec:ch4_conc}. We perform  cosmological calculations assuming $H_0 = 67.8$~km  s$^{-1}$ Mpc$^{-1}$, $\Omega_{\lambda} = 0.69$, and $\Omega_m = 0.31$ \citep{Planck2016}.  Magnitudes are reported in the AB system.

\clearpage
\section{Observations and Data Reduction}
\label{sec:ch4_obs}
GW151226 was detected on 2015 December 26 at 03:38:53 UT by a  Compact Binary Coalescence (CBC) search pipeline \citep{LIGOGW151226}. The CBC pipeline operates by matching the strain data against waveform templates and is sensitive to mergers containing NS and/or BH.  The initial sky map was generated by the {\tt BAYESTAR} algorithm and released 38 hours after the GW detection. {\tt BAYESTAR} is a Bayesian  algorithm that generates a localization sky map based on the parameter estimation from the CBC pipeline \citep{Singer+14,SingerPrice16}. The sky area contained within the initial 50\% and 90\% contours was 430 deg$^2$ and 1340 deg$^2$, respectively. A sky map generated by the {\tt LALInference} algorithm, which computes the localization using Bayesian forward-modeling  of the signal morphology \citep{Veitch+15}, was released on 2016 January 15 UT,  after our DECam observations had been concluded. The {\tt LALInference} sky map is slightly narrower than the sky map from {\tt BAYESTAR} with 50\% and 90\% contours  of 362 deg$^2$ and 1238 deg$^2$, respectively.

@@

\begin{deluxetable}{lrccccccc}
\singlespace
\tabletypesize{\footnotesize}
\tablecolumns{9}
\tablewidth{0pt}
\tablecaption{Summary of DECam Observations of GW151226 \label{tab:ch4_datatable}}
\tablehead{
\colhead{Visit}    &
\colhead{UT}       &
\colhead{$\Delta t$\tablenotemark{a}} & \colhead{$\langle$PSF$_i$$\rangle$}    & \colhead{$\langle$PSF$_z$$\rangle$}    & \colhead{$\langle$airmass$\rangle$}  & \colhead{$\langle$depth$_i\rangle$}   & \colhead{$\langle$depth$_z\rangle$}  & \colhead{$A_{\mathrm{eff}}$\tablenotemark{b}} \\
\colhead{ } &
\colhead{ } &
\colhead{(days)} &
\colhead{(arcsec)} &
\colhead{(arcsec)} &
\colhead{ } &
\colhead{(mag)} &
\colhead{(mag)} &
\colhead{(deg$^2$)}  }
\startdata
Epoch 1 & 2015-12-28.11 & 1.96 & 0.97 & 0.99 & 1.95 & 22.39 & 22.23 & 14.4 \\      %  6 hexes
& 2015-12-29.11 & 2.96 & 1.00 & 0.97 & 1.78 & 22.57 & 22.46 & 14.4 \\      %  6 hexes
Epoch 2 & 2016-01-01.06 & 5.91 & 0.95 & 0.90 & 1.57 & 21.37 & 21.06 & 28.8 \\      % 12 hexes
Epoch 3 & 2016-01-08.11 & 12.96 & 1.68 & 1.62 & 2.15 & 22.09 & 21.70 & 24.0 \\      % 10 hexes
& 2016-01-09.11 & 13.96 & 1.17 & 1.12 & 1.80 & 22.44 & 22.17 &  4.8 \\      %  2 hexes
Epoch 4 & 2016-01-18.03 & 22.88 & 1.21 & 1.20 & 1.48 & 22.00 & 22.01 & 12.0 \\      %  5 hexes
& 2016-01-19.01 & 23.86 & 1.29 & 1.25 & 1.71 & 21.86 & 21.90 & 16.8 \\ %  7 hexes
\enddata
\tablecomments{Summary of our DECam follow-up observations of GW151226. The PSF, airmass, and depth are the average values across all observations on that date. The reported depth corresponds to the mean  $5\sigma$ point source detection in the coadded search images.}
\tablenotetext{a}{Time elapsed between the GW trigger time and the time of the first image.}
\tablenotetext{b}{The effective area corresponds to 12 DECam pointings  taking into account that $\approx 20\%$ of the 3 deg$^2$ field of view of DECam is lost due to chip gaps (10\%), 3 dead CCDs (5\%, \citealt{Diehl+14}), and masked edge pixels (5\%).}
\end{deluxetable}

@@

\section{Search for an Optical Counterpart}
\label{sec:ch4_analysis}
The primary focus of our search is a fast-fading transient.  While the merger of a BBH system is not expected to produce an EM counterpart, it is informative to consider the possibility of optical emission due to the presence of some matter in the system. As a generic example, we consider the behavior of a transient such as a short gamma-ray burst (SGRB) with a typical beaming-corrected energy of $E_j \approx 10^{49}$ erg and an opening angle of $\theta_j\approx 10^\circ$ \citep{Berger2014,Fong+15}. If viewed far off-axis $(\theta_{\rm obs}\gtrsim 4\theta_j)$ the optical emission will reach peak brightness after several days, but at the distance of GW151226 ($\approx440$ Mpc, \citealt{LIGOGW151226}), the peak brightness will be $i\approx 26$ mag (see Figure 5 of \citealt{MetzgerBerger12}), well beyond our detection limit. If the source is observed moderately off-axis or on-axis $(\theta_{\rm obs} \lesssim 2\theta_j)$, then the light curve will decline throughout our observations, roughly as $F_\nu\propto t^{-1}$, and will be detectable at $i\approx$ 21--22 mag in our first observation (see Figures 3 and 4 of \citealt{MetzgerBerger12}). We can apply a similar argument to the behavior of a more isotropic (and non-relativistic) outflow given that any material ejected in a BBH merger is likely to have a low mass and the outflow will thus become optically thin early, leading to fading optical emission. Based on this line of reasoning, we search our data for steadily declining transients.

We identify relevant candidates in the data using the following selection criteria with the forced photometry from {\tt photpipe}.  Unless otherwise noted these criteria are applied to the $i$-band  data due to the greater depth in those observations.
\begin{enumerate}
\item We require non-negative or consistent with zero  (i.e., within 2$\sigma$ of zero)  $i$- and $z$-band fluxes in the difference photometry across all epochs to eliminate any sources that re-brighten in the  fourth (template) epoch. This provides an initial sample of 602 candidates.
